---
title: "Regression Modeling for Rental Property Price Prediction"
author: "Mithirendra Maniam"
date: "2023-11-15"
output: html_document
---

## 1. Introduction
Renting a property is expensive in KL and Selangor. The purpose of this project is to build a linear regression model to predict the price of a property based on its size and number of bedrooms.

## 2. Data Collection
Data was obtained from Kaggle. The URL for the dataset is given here: [Property Rental Data] (https://www.kaggle.com/datasets/ariewijaya/rent-pricing-kuala-lumpur-malaysi/download?datasetVersionNumber=5)

This dataset contains information on rent pricing surrounding Kuala Lumpur and Selangor region, Malaysia. The information was scraped from mudah.my.

There are 13 features with one unique ids (ads_id) and one target feature (monthly_rent)

- <code>ads_id</code>: the listing ids (unique)
- <code>prop_name</code>: name of the building/ property
- <code>completion_year</code>: completion/ established year of the property
- <code>monthly_rent</code>: monthly rent in ringgit malaysia (RM)
- <code>location</code>: property location in Kuala Lumpur region
- <code>property_type</code>:property type such as apartment, condominium, flat, duplex, studio, etc
- <code>rooms</code>: number of rooms in the unit
- <code>parking</code>: number of parking space for the unit
- <code>bathroom</code>: number of bathrooms in the unit
- <code>size</code>: total area of the unit in square feet
- <code>furnished</code>: furnishing status of the unit (fully, partial, non-furnished)
- <code>facilities</code>: main facilities available
- <code>additional_facilities</code>: additional facilities (proximity to attraction area, mall, school, shopping, railways, etc)

Reading of data into R
```{r}
library(dplyr)

# Reading data from csv file
prop_data <- read.csv("mudah-apartment-kl-selangor.csv")
```




## 3. Data Cleaning

Check the dimensions of the data set.

```{r}
# Check the number of observations and columns in dataset
dim(prop_data)
```

Next we check the structure of the dataset.

```{r}
# Looking at structure of dataset
str(prop_data)
```

We are looking for 3 variables to perform the prediction, namely <code>monthly_rent</code>, <code>size</code> and <code>rooms</code>. From the structure of the dataset above, it can be seen that all 3 variables are character classes. We should subset the columns in a new variable and convert this character variables into numeric variables so that we can run a regression model.

Also added the <code>property type</code> column to be able to choose the type of property, to allow for more accurate prediction.

```{r}
# Subset a few columns to create a new subset dataset
subprop_data <- prop_data[ , c(4,6,7,10)]
str(subprop_data)
```


### 3.1 Property Size Data

Below are the steps to prepare the Size column.

```{r, results='hide'}
# Change column from character to numeric
library(stringr)
subprop_data$size <- str_replace(subprop_data$size, " sq.ft.", "")
subprop_data$size <- as.numeric(subprop_data$size)
str(subprop_data)
```


### 3.2 Property Room Number

Below are the steps to prepare the Room column.

```{r, results='hide'}
# Change column from character to numeric
subprop_data$rooms <- as.numeric(subprop_data$rooms)
str(subprop_data)
```

### 3.3 Property Monthly Rent

Below are the steps to prepare the monthly_rent column.

```{r, results='hide'}
# Change column from character to numeric
subprop_data$monthly_rent <- str_replace(subprop_data$monthly_rent, " per month", "")
subprop_data$monthly_rent <- str_replace(subprop_data$monthly_rent, "RM ", "")
subprop_data$monthly_rent <- str_replace(subprop_data$monthly_rent, " ", "")
```

Convert monthly_rent column to numeric values.

```{r}
# Convert monthly-rent column to numeric type
subprop_data$monthly_rent <- as.numeric(subprop_data$monthly_rent)
str(subprop_data)
```

### 3.4 Property Type

Converting the Property Type column into factors and finding unique levels within column.

```{r}

# Remove data with weird property_type description
subprop_data[subprop_data$property_type == "Condo / Services residence / Penthouse / Townhouse", ]
subprop_data <- subprop_data[-3669, ]

# Change column from character to numeric
subprop_data$property_type <- factor(subprop_data$property_type)
levels(subprop_data$property_type)

```

## 4. Exploratory Data Analysis

```{r}
# Remove NAs
subprop_data <- na.omit(subprop_data)

# Drop unused levels after removing NAs
subprop_data <- droplevels(subprop_data)

# Summary statistics
summary(subprop_data$size)
summary(subprop_data$rooms)
summary(subprop_data$monthly_rent)

# Checking for outliers
tail(sort(subprop_data$size),20)
tail(sort(subprop_data$rooms),20)
tail(sort(subprop_data$monthly_rent),50)


# Checking and deleting size column outliers
subprop_data[subprop_data$size > 40000,]
subprop_data <- subprop_data[-c(1715,2623),]

# Summary statistics
summary(subprop_data$size)

# Checking and deleting monthly_rent column > 50000
subprop_data <- subprop_data[subprop_data$monthly_rent < 20000,]

# Check the distribution of monthly rent
plot(subprop_data$size, subprop_data$monthly_rent,
     xlim = c(0,2000),
     xlab = "Size in sq feet",
     ylab = "Monthly Rent (in RM)",
     pch = 19,
     col = subprop_data$property_type,
     main = "Monthly Rent by Size of Property (in Sq Feet)")
legend("topleft",
       legend = levels(subprop_data$property_type),
       pch = 19,
       col = subprop_data$property_type)

plot(subprop_data$room, subprop_data$monthly_rent,
     xlab = "Number of rooms",
     ylab = "Monthly Rent (in RM)",
     pch = 19,
     col = subprop_data$property_type,
     main = "Monthly Rent by Number of Rooms in Property")
legend("topleft",
       legend = levels(subprop_data$property_type),
       pch = 19,
       col = subprop_data$property_type)

# Check mean for each property type
tapply(subprop_data$monthly_rent, subprop_data$property_type, mean)

```

## 5. Data Modeling and Prediction

Building a linear regression model using size and rooms to predict monthly rent.

```{r}
# Build a linear regression model
model <- lm(monthly_rent ~ rooms + size, data=subprop_data)
summary(model)
```

## 6. Outcome - Predicting Price based on input of size and rooms

```{r}
# Function for predicting property rental
rental_predictor <- function(type, size, rooms) {
  subprop_type <- subprop_data[subprop_data$property_type == type,  ]
  model <- lm(monthly_rent ~ rooms + size, data=subprop_type)
  newdata <- data.frame(size, rooms)
  prediction <- predict(model, newdata)
  print(paste("The predicted rental price for a", type, "property with", rooms, "rooms and a size of", size,"sq feet is: ~RM", round(prediction,0)))
}
```

```{r}
rental_predictor("Apartment",1200, 3)
rental_predictor("Condominium",1000,3)
rental_predictor("Studio",400,1)
```
```

